Blum axioms and nondeterministic computation of functions

### 2016-03-22 paper.tex
paper.tex: Add introductory section about the quality standards.

Most extensions to this basic model,
like the use of several tracks, or several tapes, or multidimensional tapes,
can be easily incorporated in the definition of polynomial-time computable.
Except nondeterminism.%
\footnote{
    Note also all the other extensions can be simulated by the original model
    suffering at most a polynomial slowdown,
    even in the context of decision problems.
    This is seems not to be the case with nondeterminism.
}
We hit a wall right in the start:
how does a nondeterministic machine computes a function in the first place?

This paper presents several attempts to establish
how a nondeterministic Turing machine could compute a function,
and to extend the concept of ``polynomial-time computable'' under each definition.
Sections \ref{sec:godel_numberings} and~\ref{sec:blum-axioms} sets reasonableness criteria
to both the definition of function computation and complexity of the computation.
Section~\ref{sec:nondeterministic_computation_of_functions}
presents and criticizes the several definitions.

>>>>

Most extensions to this basic model,
like the use of several tracks, or several tapes, or multidimensional tapes,
can be easily incorporated in the definition of polynomial-time computable.
Except nondeterminism.
We hit a wall right in the start:
how does a nondeterministic machine computes a function in the first place?

This paper presents several attempts to establish
how a nondeterministic Turing machine could compute a function,
and to extend the concept of ``polynomial-time computable'' under each definition.
Section~\ref{sec:quality-standards} sets reasonableness criteria
to both the definition of function computation and complexity of the computation.
Section~\ref{sec:nondeterministic_computation_of_functions}
presents and criticizes the several definitions.

\section{Quality Standards: Blum Axioms}
\label{sec:quality-standards}

The definition of non-deterministic computation of functions we are trying to make
must encompass both the ``computation'' part
and the ``non-deterministic'' part.

In the context of deciders,
allowing doubly-infinite tapes and multiple tapes,
although simplifies the design and increases the efficiency of the machine,
does not expand the class of languages that can be decided or recognized by the machine.
That is, both models are \emph{computationally equivalent}.
Therefore,
our definition must produce an equivalent model
--- every function that can be computed by deterministic means
must be computable under our new definition and vice-versa.
This is the ``computation'' part.

But nondeterministic deciders,
although computationally equivalent to deterministic ones,
are apparently much faster.
That is, while we have linear-time non-deterministic algorithms for Boolean satisfiability,
the best known deterministic algorithms are still exponential in the worst case.
Nondeterminism seems special in this sense;
although other extensions (multiple tapes and doubly-infinite tapes)
can be simulated under just a polynomial slowdown,
the simulation of nondeterminism seems to require exponential slowdown.%
\footnote{
    Simulating nondeterminism under a polynomial slowdown would show that $\P = \NP$;
    showing that the slowdown must be exponential would show that $\P \neq \NP$.
    Although most researchers expect the later to be the case~\cite[p.~54]{Gasarch2012},
    as Papadimitriou noted~\cite[p.~412]{Papadimitriou1994},
    in the absence of a proof that $\P \neq \NP$
    we should not be too emphatic in stating the simulation \emph{require}
    (as opposed to \emph{seemingly require}) exponential slowdown.
}
Therefore,
our definition must exhibit this apparent exponential speedup
when computing, for instance, the characteristic function of $\SAT$.
This is the ``non-deterministic'' part.

Sections \ref{sec:godel_numberings} and~\ref{sec:blum-axioms}
refine this concept.
